The \kepler Smear Campaign I: An Asteroseismic Catalogue of Bright Red Giants

### 2018-12-05 paper/ms.tex
jofre changes

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{longtable}

@@

A major outcome of the \kepler asteroseismology programme is a legacy sample of extremely well characterized stars that can serve as benchmarks for future work \citep{keplerlegacy1,keplerlegacy2,silvaplanetages,2016MNRAS.456.2183D}. Asteroseismological studies with \kepler complement other probes of stellar physics well: for example, the APOKASC sample of~1916 spectroscopically- and asteroseismically-characterized red giant stars \citep{2014ApJS..215...19P}. For this APOKASC sample, \citet{hawkinsapogee} have been able to extract precise elemental abundances by fitting spectroscopic data with \logg and \teff fixed to asteroseismically-determined values. It is necessary to calibrate such a study against benchmark stars with very precisely-determined parameters, which in practice requires nearby bright stars that are amenable to very high signal-to-noise spectroscopy plus asteroseismology \citep{creeveybenchmark}, parallaxes \citep{hawkinsbenchmarks}, and/or interferometry \citep{casagrandebenchmark,creeveybenchmark2}. This is especially important in the context of the \gaia mission \citep{gaia}, which has recently put out its second data release of 1,692,919,135 sources, including 1,331,909,727 with parallaxes \citep{gaiadr2}. These data will form the basis of many large surveys and it is vital that they are calibrated correctly. To this end, 34~FGK stars including both giants and dwarfs have been chosen as \gaia-ESO benchmark stars for which metallicities \citep{gaiabenchmark1}, effective temperatures and asteroseismic surface gravities \citep{gaiabenchmark3}, and relative abundances of $\alpha$ and iron-peak elements \citep{gaiabenchmark4} have been determined. This includes only four main-sequence stars much cooler than the Sun, due to the paucity of such stars with asteroseismology. This has been accompanied by the release of high resolution spectra \citep{gaiabenchmark2} and formed the basis of extensions to lower metallicities \citep{gaiabenchmark5}, stellar twin studies \citep{gaiabenchmarktwins} and comparisons of stellar abundance determination pipelines \citep{gaiabenchmarkabundances}. Furthermore, by combining asteroseismology with optical interferometry, it has been possible to determine fundamental parameters of main-sequence and giant stars with unprecedented precision \citep{huber12,thetacygwhite,white15}.

Brighter \kepler stars are therefore ideal benchmark targets, since photometry can be most easily complemented by \gaia parallaxes, interferometric diameters, and high resolution spectroscopy.  
Unfortunately, the \kepler field was deliberately placed to minimize overall the number of extremely bright stars on the detectors, so that only a dozen stars brighter than~6th~magnitude landed on silicon \citep{2010ApJ...713L..79K}. This was because stars brighter than $Kp \sim 11$ saturate the CCD detectors, with their flux distributed along a bleed column and rendering those pixels otherwise unusable. Furthermore, due to the limited availablility of bandwidth to download data from the spacecraft, only $\sim 5.7\%$ of pixels on the \kepler detectors are actually downloaded in any one Quarter \citep{2010ApJ...713L..87J}. The result of these two target selection constraints is that photometry was obtained for most of the mission for only~35 stars brighter than $Kp<7$ in the \kepler field, while~17 targets in this range were observed for less than half the mission and~29 targets brighter than this threshold were entirely ignored. The availability of \kepler data remains significantly incomplete down to fainter magnitudes, and in this work we consider $Kp=9$ to be an arbitrary cutoff for bright stars of interest. In the \ktwo mission \citep{k2early}, very saturated stars have been observed with `halo photometry' using unsaturated pixels in a specially-determined region around bright stars, including the Pleiades \citep{halo}, Aldebaran \citep{aldebaran}, and $\rho$~Leonis \citep{rholeo}. Unfortunately, in the legacy \kepler sample, photometry of such saturated stars was rarely attempted.

@@

We have obtained high-resolution spectroscopy with TRES for~63 stars, mainly giants, in order to constrain stellar parameters and elemental abundances. Operating with spectral resolving power $R=44 000$, we have obtained spectra with a mean signal-to-noise ratio (SNRs) of $\sim 100$  per resolution element. Although this resolution and SNR are sufficient for an exploratory study, a more detailed analysis will require APOGEE or a similar instrument to obtain greater spectral coverage. From this observing run we have~33 unique targets with seismic \logg and spectra, the same as in the \gaia-ESO benchmark set and a significant addition to the ensemble of bright red giants with asteroseismic parameter determinations. %Due to observing constraints, we were unable to obtain spectra for BD+42~3150, BD+48~2904, HD~176209, HD~182354, HD~189636AB, or HD~189750. % figure out this discrepancy?

To derive stellar parameters from our TRES spectra, we initially ran the Stellar Parameter Classification code \citep[SPC:][]{spc} to determine \teff and \logg, using the SPC \teff to inform the asteroseismic estimation of \logg from \numax. For deriving abundances, \teff was fixed from the results of an initial SPC fit, while \logg was fixed to the seismic values. The other stellar atmospheric parameters including the microturbulent velocity ($v_{\text{mic}}$), broadening (convolution by $V_{\text{mac}}$, $v_{\sin{i}}$ and the instrumental line profile), as well as [Fe/H] and chemical abundances for 20 chemical species were derived using the Brussels Automatic Code for Characterizing High accUracy Spectra \citep[BACCHUS:][]{bacchus}. The results from this calculation are displayed in Table~\ref{stellar_props}. BACCHUS uses an interpolation scheme through a grid of MARCS model atmospheres \citep{Gustafsson2008} in combination with TURBOSPECTRUM \citep{Alvarez1998, Plez2012}. For the calculation of synthetic spectra, atomic line information has been taken from the fifth version of the Gaia-ESO linelist (Heiter et al., in preparation). Additionally, we used the molecular species for CH \citep{Masseron2014}, CN, NH, OH, MgH  C$_{2}$ (T. Masseron, private communication). The SiH molecular information was adopted from the Kurucz linelists and the information for TiO, ZrO, FeH, CaH from B. Plez (private communication).

@@

To achieve the most precise abundances we have derived them both with and without a line-by-line differential approach with respect to Arcturus ($\alpha$~Bo\"{o}tis), using the method described by \citet{gaiabenchmark4} and the Arcturus abundances from \citep{hawkinsapogee}. Choosing this method means we do not derive the abundances for neutron capture  elements (e.g. Sr, Y, Zr, Ba, La, Nd, Eu) in a differential way because there are no estimated values for these elements at the appropriate benchmark parameters of Arcturus. For these elements we instead derived the chemical abundances in an absolute way, where the solar abundances of \cite{Asplund2005} were assumed. The uncertainty in the abundances are taken as the line-by-line dispersion, or assumed to be 0.10 when just one line is available. No abundances for oxygen could be reliably derived for any of the stars in our spectroscopic sample by either method.

@@

We compare the solutions found using this set of fitting parameter with those found using only spectroscopic input in Figure~\ref{masses}, and we find that the change in median stellar parameters between the two results is small for all analyzed red giants. However, in the comparison plot it becomes clear that adding the asteroseismic constraints to the fit narrows down the parameter space and thus decreases the uncertainties.

@@

One of the primary ways to determine, in a chemical sense, whether the stars in our sample are drawn from the Galactic disk(s), bulge or halo, is with the ratio of their $\alpha$-elements to Fe. The $\alpha$ elements are formed during He burning (e.g. Mg, Ti, Si, Ca) and largely dispersed into the interstellar medium through Type~II supernovae (SNII) \citep{Matteucci2001}. It has been shown by many studies \citep[e.g.][and references therein]{Edvardsson1993, Adibekyan2012, Feltzing2013, Bensby2014}, that the Galactic thick and thin disks separate in the $\alpha$ elements, where the thick disk is enhanced in [Mg, Si, Ca, Ti/Fe] compared to the Galactic thin disk at a given metallicity. In Fig.~\ref{alphael}, we display the [Mg/Fe] abundance ratio as a function of [Fe/H] for our stars (black circles), compared\footnote{There may to be systematics between our [X/Fe] abundance scale and those of our comparison samples.} to representative thick and thin disk stars from \cite[open red squares]{Bensby2014} and \cite[open orange triangles]{Adibekyan2012}.

@@

The coherent g-mode pulsations in samples of B, A, and F stars observed by \kepler previously showed these stars to be near-rigid rotators \citep{kurtz2014,saio15,triana15,vanreeth15,vanreeth16,vanreeth18,murphy16,schmid16,moravveji16,ouazzani17,papics17,aerts17,szewczuk18,2018arXiv180907779A}. These studies cover about 70 stars so far. However, the vast majority of intermediate-mass stars observed by Kepler have yet to be subjected to in-depth asteroseismic analyses and modelling of their interior properties. One of the valuable outputs of our current work includes the reduced light curves of several early-B stars, which were only scarcely targeted in the nominal \kepler mission. The few that were monitored did not reveal suitable oscillation frequency patterns to achieve a unique mode identification, which is a requirement to perform asteroseismic modelling. The investigation of pulsation modes in high-mass stars using high-quality \kepler smear data combined with high-precision spectroscopy to identify the modes \citep[Chapter 6]{aertsbook} is an exciting prospect for asteroseismology, as the interior physics of these stars are largely unknown, yet they play a pivotal role in stellar and galactic evolution. The in-depth asteroseismic analysis of the smear data for the~B~stars requires additional ground-based follow-up spectroscopy and will be the subject of future work.

@@

This work was performed in part under contract with the Jet Propulsion Laboratory (JPL) funded by NASA through the Sagan Fellowship Program executed by the NASA Exoplanet Science Institute. B.P. also acknowledges support from Balliol College and the Clarendon Fund. D.H. acknowledges support by the National Aeronautics and Space Administration through the K2 Guest Observer Program (NNX17AF76G, 80NSSC18K0362) and the National Science Foundation (AST-1717000). DWL acknowledges partial support from the Kepler Extended Mission under NASA Cooperative Agreement NNX13AB58A with the Smithsonian Astrophysical Observatory. The research leading to these results has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement No670519: MAMSIE). V.S.A. acknowledges support the Independent Research Fund Denmark (Research grant 7027-00096B). Funding for the Stellar Astrophysics Centre is provided by The Danish National Research Foundation (Grant agreement No.~DNRF106).

>>>>

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{longtable}
\usepackage{ulem}

@@

A major outcome of the \kepler asteroseismology programme is a legacy sample of extremely well characterized stars that can serve as benchmarks for future work \citep{keplerlegacy1,keplerlegacy2,silvaplanetages,2016MNRAS.456.2183D}. Asteroseismological studies with \kepler complement other probes of stellar physics well: for example, the APOKASC sample of~1916 spectroscopically- and asteroseismically-characterized red giant stars \citep{2014ApJS..215...19P}. For this APOKASC sample, \citet{hawkinsapogee} have been able to extract precise elemental abundances by fitting spectroscopic data with \logg and \teff fixed to asteroseismically-determined values. It is necessary to calibrate such a study against benchmark stars with very precisely-determined parameters, which in practice requires nearby bright stars that are amenable to very high signal-to-noise spectroscopy plus asteroseismology \citep{creeveybenchmark}, parallaxes \citep{hawkinsbenchmarks}, and/or interferometry \citep{casagrandebenchmark,creeveybenchmark2}. This is especially important in the context of the \gaia mission \citep{gaia}, which has recently put out its second data release of 1,692,919,135 sources, including 1,331,909,727 with parallaxes \citep{gaiadr2}. These data will form the basis of many large surveys and it is vital that they are calibrated correctly. To this end, 36~FGK stars including both giants and dwarfs have been chosen as \gaia benchmark stars for which metallicities \citep{gaiabenchmark1,2018RNAAS...2c.152J}, effective temperatures and asteroseismic surface gravities \citep{gaiabenchmark3}, and relative abundances of $\alpha$ and iron-peak elements \citep{gaiabenchmark4} have been determined. This includes only four main-sequence stars much cooler than the Sun, due to the paucity of such stars with asteroseismology. This has been accompanied by the release of high resolution spectra \citep{gaiabenchmark2} and formed the basis of extensions to lower metallicities \citep{gaiabenchmark5}, stellar twin studies \citep{gaiabenchmarktwins} and comparisons of stellar abundance determination pipelines \citep{gaiabenchmarkabundances}. Furthermore, by combining asteroseismology with optical interferometry, it has been possible to determine fundamental parameters of main-sequence and giant stars with unprecedented precision \citep{huber12,thetacygwhite,white15}.

Brighter \kepler stars are therefore ideal benchmark targets, since photometry can be most easily complemented by \gaia parallaxes, interferometric diameters, and high resolution spectroscopy.  
Unfortunately, the \kepler field was deliberately placed to minimize overall the number of extremely bright stars on the detectors, so that only a dozen stars brighter than~6th~magnitude landed on silicon \citep{2010ApJ...713L..79K}. This was because stars brighter than $Kp \sim 11$ saturate the CCD detectors, with their flux distributed along a bleed column and rendering those pixels otherwise unusable. Furthermore, due to the limited availablility of bandwidth to download data from the spacecraft, only $\sim 5.7\%$ of pixels on the \kepler detectors are actually downloaded in any one Quarter \citep{2010ApJ...713L..87J}. The result of these two target selection constraints is that photometry was obtained for most of the mission for only~35 stars brighter than $Kp<7$ in the \kepler field, while~17 targets in this range were observed for less than half the mission and~29 targets brighter than this threshold were entirely ignored. The availability of \kepler data remains significantly incomplete down to fainter magnitudes, and in this work we consider $Kp=9$ to be an arbitrary cutoff for bright stars of interest. In the \ktwo mission \citep{k2early}, very saturated stars have been observed with `halo photometry' using unsaturated pixels in a specially-determined region around bright stars, including the Pleiades \citep{halo}, Aldebaran \citep{aldebaran}, and $\rho$~Leonis \citep{rholeo}. Unfortunately, in the four-year \kepler sample, photometry of such saturated stars was rarely attempted.

@@

We have obtained high-resolution spectroscopy with TRES for~63 stars, mainly giants, in order to constrain stellar parameters and elemental abundances. Operating with spectral resolving power $R=44 000$, we have obtained spectra with a mean signal-to-noise ratio (SNRs) of $\sim 100$  per resolution element. Although this resolution and SNR are sufficient for an exploratory study, a more detailed analysis will require APOGEE or a similar instrument to obtain greater spectral coverage. From this observing run we have~33 unique targets with seismic \logg and spectra, comparable to the 36 of the \gaia benchmark set \citep{2018RNAAS...2c.152J} and a significant addition to the ensemble of bright red giants with asteroseismic parameter determinations. %Due to observing constraints, we were unable to obtain spectra for BD+42~3150, BD+48~2904, HD~176209, HD~182354, HD~189636AB, or HD~189750. % figure out this discrepancy?

To derive stellar parameters from our TRES spectra, we initially ran the Stellar Parameter Classification code \citep[SPC:][]{spc} to determine \teff and \logg, using the SPC \teff to inform the asteroseismic estimation of \logg from \numax. For deriving abundances, \teff was fixed from the results of an initial SPC fit, while \logg was fixed to the seismic values. The other stellar atmospheric parameters including the microturbulent velocity ($v_{\text{mic}}$), broadening (convolution by $v_{\text{mac}}$, $v_{\sin{i}}$ and the instrumental line profile), as well as [Fe/H] and chemical abundances for 20 chemical species were derived using the Brussels Automatic Code for Characterizing High accUracy Spectra \citep[BACCHUS:][]{bacchus}. The results from this calculation are displayed in Table~\ref{stellar_props}. BACCHUS uses an interpolation scheme through a grid of MARCS model atmospheres \citep{Gustafsson2008} in combination with TURBOSPECTRUM \citep{Alvarez1998, Plez2012}. For the calculation of synthetic spectra, atomic line information has been taken from the fifth version of the Gaia-ESO linelist (Heiter et al., in preparation). Additionally, we used the molecular species for CH \citep{Masseron2014}, CN, NH, OH, MgH  C$_{2}$ (T. Masseron, private communication). The SiH molecular information was adopted from the Kurucz linelists and the information for TiO, ZrO, FeH, CaH from B. Plez (private communication).

@@

To achieve the most precise abundances we have derived them both with and without a line-by-line differential approach with respect to Arcturus ($\alpha$~Bo\"{o}tis), using the method described by \citet{hawkinsapogee} and the Arcturus abundances from \citep{gaiabenchmark4}. Choosing this method means we do not derive the abundances for neutron capture  elements (e.g. Sr, Y, Zr, Ba, La, Nd, Eu) in a differential way because there are no estimated values for these elements at the appropriate benchmark parameters of Arcturus. For these elements we instead derived the chemical abundances in an absolute way, where the solar abundances of \cite{Asplund2005} were assumed. The uncertainty in the abundances are taken as the line-by-line dispersion, or assumed to be 0.10 when just one line is available. No abundances for oxygen could be reliably derived for any of the stars in our spectroscopic sample by either method.

@@

We compare the solutions found using this set of fitting parameter with those found using only spectroscopic input in Figure~\ref{masses}, and we find that the change in median stellar parameters between the two results is small for all analyzed red giants. However, in the comparison plot it becomes clear that adding the asteroseismic constraints to the fit reduces the posterior uncertainty in stellar mass.

@@

One of the primary ways to determine, in a chemical sense, whether the stars in our sample are drawn from the Galactic disk(s), bulge or halo, is with the ratio of their $\alpha$-elements to Fe. The $\alpha$ elements are formed after He burning (e.g. Mg, Ti, Si, Ca) and largely dispersed into the interstellar medium through Type~II supernovae (SNII) \citep{Matteucci2001}. It has been shown by many studies \citep[e.g.][and references therein]{Edvardsson1993, Adibekyan2012, Feltzing2013, Bensby2014}, that the Galactic thick and thin disks separate in the $\alpha$ elements, where the thick disk is enhanced in [Mg, Si, Ca, Ti/Fe] compared to the Galactic thin disk at a given metallicity. In Fig.~\ref{alphael}, we display the [Mg/Fe] abundance ratio as a function of [Fe/H] for our stars (black circles), compared\footnote{There may to be systematics between our [X/Fe] abundance scale and those of our comparison samples.} to representative thick and thin disk stars from \cite[open red squares]{Bensby2014} and \cite[open orange triangles]{Adibekyan2012}.

@@

The coherent g-mode pulsations in samples of B, A, and F stars observed by \kepler previously showed these stars to be near-rigid rotators \citep{kurtz2014,saio15,triana15,vanreeth15,vanreeth16,vanreeth18,murphy16,schmid16,moravveji16,ouazzani17,papics17,aerts17,szewczuk18,2018arXiv180907779A}. These studies cover about 70 stars so far. However, the vast majority of intermediate-mass stars observed by Kepler have yet to be subjected to in-depth asteroseismic analyses and modelling of their interior properties. One of the valuable outputs of our current work includes the reduced light curves of several early-B stars, which were only scarcely targeted in the nominal \kepler mission. The few that were monitored did not reveal suitable oscillation frequency patterns to achieve a unique mode identification, which is a requirement to perform asteroseismic modelling. The investigation of pulsation modes in high-mass stars using high-quality \kepler smear data combined with high-precision spectroscopy to identify the modes \citep[Chapter 6]{aertsbook} is an exciting prospect for asteroseismology, as the interior physics of these stars are largely unknown, yet they play a pivotal role in stellar and galactic evolution. The in-depth asteroseismic analysis of the smear data for the~B~stars will be the subject of future work.

@@

This work was performed in part under contract with the Jet Propulsion Laboratory (JPL) funded by NASA through the Sagan Fellowship Program executed by the NASA Exoplanet Science Institute. B.P. also acknowledges support from Balliol College and the Clarendon Fund. D.H. acknowledges support by the National Aeronautics and Space Administration through the K2 Guest Observer Program (NNX17AF76G, 80NSSC18K0362) and the National Science Foundation (AST-1717000). DWL acknowledges partial support from the Kepler Extended Mission under NASA Cooperative Agreement NNX13AB58A with the Smithsonian Astrophysical Observatory. The research leading to these results has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement No670519: MAMSIE). V.S.A. acknowledges support the Independent Research Fund Denmark (Research grant 7027-00096B). Funding for the Stellar Astrophysics Centre is provided by The Danish National Research Foundation (Grant agreement No.~DNRF106). PJ acknowledges FONDECYT Iniciaci\'on grant 11170174.
